---
title: "COMPASS LE OWC Root Exclusion: Porewater Nutrients"
author: "Apr-May 2024 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

## Run Information 
```{r run information, include=TRUE}
#set the run date & user name 
  run_date <- "08/20/2025"
  sample_year <- "2024"
  sample_month <- "Apr-May"
  user <- "Roberta Peixoto"
  QAQC_by <- "Zoe Read"

#identify the files you want to read in 
  #read in as a list to accommodate multiple runs in a month
  data_files <- c("Raw Data/Seal_COMPASS_LE_OWC_RootExclusion_NUTR_Apr-May2024.csv")

# Define the file path for QAQC log file - NO Need to change just check year 
  file_path <- "Raw Data/SEAL_COMPASS_LE_OWC_RootExclusion_QAQC_Log_2025.csv"
  final_path <- "Processed Data/COMPASS_LE_OWC_RootExclusion_Nutrients_2024_Apr-May2024.csv"

#record any notes about the run or anything other info here: 
  run_notes <- "The last CCV check std for Phosphate is very high. 
  All other check standards are too low. " #Add notes from QAQC
  
#Set up file path for metadata 
  #downloaded metadata csv - downloaded from Google drive as csv for this year
  Raw_Metadata = "Raw Data/COMPASS_OWC_Root_Exclusion_Nutrient_SampleLog.csv"

  cat(run_notes)
```

##Setup
```{r setup, include=FALSE}

#let you know which section you are in 
cat("Setup")

#a link to the Gitbook or whatever protocol you are using for this analysis 


#load/install packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  purrr,
  ggplot2,
  ggpubr,
  tidyr,
  tidyverse,
  lubridate,
  tinytex,
  stringr,
  writexl,
  readr,
  readxl,
  purrr,
  tinytex,
  broom, 
  patchwork)

#Coefficients / constants that are needed for calculations 
N_mw <- 14.0067    # molecular weight of N 
P_mw <- 30.973762  # molecular weight of P 
S_mw <- 32.06      # molecular weight of S
Con1 <- 1000       # conversion factor value
Con2 <- 1000000    # conversion factor value 

#Detection limit and top standards for flagging 
####Need to determine detection limits for these analytes for this SEAL####
NOx_dl <- 0.025  #mgL
NH3_dl <- 0.02  #mgL
PO4_dl <- 0.003 #mgL
NO2_dl <- 0
SO4_dl <- 0

NOx_top <- 1    #ug/L
NH3_top <- 2    #ug/L
PO4_top <- 1    #ug/L
NO2_top <- 40   #ug/L
SO4_top <- 200  #ug/L

#Check standard expected concentrations 
##We set 10% variability as acceptable, with values expected to be either the average between the CCB (blank) and the top standard for each variable, or half of the top standard.
NOx_CCV <- NOx_top/2 
NH3_CCV <- NH3_top/2 
PO4_CCV <- PO4_top/2 
SO4_CCV <- NO2_top/2
NO2_CCV <- SO4_top/2

# #Spike Concentrations 
# NOx_Spk <- 0.5 #ppm or mg/L
# NH3_Spk <- 1
# PO4_Spk <- 0.152 

#expected ranges for sample concentrations used for flags 
r2_cutoff = 0.990
chk_flag = 0.25
chk_conc_flag = 10 #cutoff level for percent difference of check stds vs. expected concentration, 10% for OWC LE Root Exclusion experiment
chks_flag = 60
rep_flag = 25 
#^this is a 25% error between samples
#blank_flag is calculated based on samples later in this code

# #pe check Concentrations
# NH3_pe <- 1.034
# NOx_pe <- 1.51
# PO4_pe <- 0.824

```

##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in the raw metadata file 
raw_metadata <- read.csv(Raw_Metadata)

head(raw_metadata)
# 1_1_R34_4192024_ch1_trC_cm

##Format Date column 
raw_metadata$Date_formatted <- paste0(
  raw_metadata$Year,
  sprintf("%02d", raw_metadata$Month),
  sprintf("%02d", raw_metadata$Day)
)

#Create NUTR IDs from what was collected for comparison later
raw_metadata <- raw_metadata %>%
  mutate(NUTR_ID = paste(Box,
                         Vial_number, 
                         Vial_ID, 
                         Date_formatted, 
                         paste0("ch", Chamber_ID), 
                         paste0("tr", Treatment),
                         paste0(Depth_cm, "cm"),
                        sep = "_"), 
                              NA) 

#Take out the columns and rows that are not relevant
nutr_metadata <- raw_metadata %>%
  select(-"NA") %>%
  # only keep specific columns
  filter(!is.na(NUTR_ID) & NUTR_ID != "")  # remove missing/blank NUTR_ID rows

```

\newpage

##Import Data & Clean
```{r Import Data, include=FALSE}
cat("Import Data")

#set file path for data - in run info chunk 
path <- ("file path")

#Read in Raw Data 
data <- read.csv(data_files)

data$Dilution = (data$Man.Dil.Factor*data$Auto.Dil.Factor)
data$Run_number = 1

# Rename columns
df_all <- data %>%
  select(Sample_Name = Sample.ID,
         Run_Number = Run_number,
         Conc = Results,
         Absorbance = Absorbance,
         Dilution = Dilution,
         Unit = Units,
         Test = Test,
         Run_Time = Date.and.Time) 

#Checking column headers
head(df_all)


```

##Assessing standard Curves
```{r Pull out Standard Curves, echo=FALSE}

####Added SO4 and NO2 std curve, NOX curve is now linear and SO4 curve is a third-order polynomial####

#Pull out standards 
stds <- df_all %>%
  filter(str_detect(Sample_Name, "Standard | NO2"))

#Making standards dataframe based on the test
stds_NOx <- stds[stds$Test == "Nitrate (NOx)", ]
stds_NH3 <- stds[stds$Test == "Ammonia (PPS)", ]
stds_PO4 <- stds[stds$Test == "Phosphate ", ]
stds_NO2 <- stds[stds$Test == "Nitrite", ]
stds_SO4 <- stds[stds$Test == "Sulfate", ]

#Inputting Standard Values Based on Protocol (LINK PROTOCOL)
stds_NOx <- stds_NOx %>%
  mutate(`Conc` = case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard 90" ~ 0.0111,
    Sample_Name == "Standard 91" ~ 0.0222,
    Sample_Name == "Standard 92" ~ 0.0444,
    Sample_Name == "Standard 93" ~ 0.0500,
    Sample_Name == "Standard 94" ~ 0.1,
    Sample_Name == "Standard 95" ~ 0.2,
    Sample_Name == "Standard 96" ~ 0.3,
    Sample_Name == "Standard 97" ~ 0.5,
    Sample_Name == "Standard 98" ~ 0.75,
    Sample_Name == "Standard 99" ~ 1,
    TRUE ~ `Conc`),  # leave unchanged if no match
    Run_Date = format(lubridate::ymd_hms(Run_Time), "%m-%d-%Y")) %>%
  filter(!Sample_Name %in% c("Nitrate Standard" , "Nitrite Standard"))

#ID x and y
x1 <- stds_NOx$Absorbance
y1 <- stds_NOx$Conc

stds_NH3 <- stds_NH3 %>%
  mutate(`Conc` = case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard 90" ~ 0.04,
    Sample_Name == "Standard 91" ~ 0.08,
    Sample_Name == "Standard 92" ~ 0.12,
    Sample_Name == "Standard 93" ~ 0.16,
    Sample_Name == "Standard 94" ~ 0.2,
    Sample_Name == "Standard 95" ~ 0.4,
    Sample_Name == "Standard 96" ~ 0.6,
    Sample_Name == "Standard 97" ~ 1,
    Sample_Name == "Standard 98" ~ 1.5,
    Sample_Name == "Standard 99" ~ 2,
    TRUE ~ `Conc`),  # leave unchanged if no match
    Run_Date = format(lubridate::ymd_hms(Run_Time), "%m-%d-%Y"))

# identifying x and y
x2 <- stds_NH3$Absorbance
y2 <- stds_NH3$Conc

stds_PO4 <- stds_PO4 %>%
  mutate(`Conc` = case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard 90" ~ 0.02,
    Sample_Name == "Standard 91" ~ 0.04,# RBP changed from 0.05 to 0.04 in 2/18/2026
    Sample_Name == "Standard 92" ~ 0.06,
    Sample_Name == "Standard 93" ~ 0.1,
    Sample_Name == "Standard 94" ~ 0.2,
    Sample_Name == "Standard 95" ~ 0.3,
    Sample_Name == "Standard 96" ~ 0.4,
    Sample_Name == "Standard 97" ~ 0.6,
    Sample_Name == "Standard 98" ~ 0.8,
    Sample_Name == "Standard 99" ~ 1,
    TRUE ~ `Conc`),  # leave unchanged if no match
    Run_Date = format(lubridate::ymd_hms(Run_Time), "%m-%d-%Y"))

#ID x and y
x3 <- stds_PO4$Absorbance
y3 <- stds_PO4$Conc

stds_SO4 <- stds_SO4 %>%
  mutate(`Conc` = case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard 90" ~ 4,
    Sample_Name == "Standard 91" ~ 12,
    Sample_Name == "Standard 92" ~ 16,
    Sample_Name == "Standard 93" ~ 20,
    Sample_Name == "Standard 94" ~ 24,
    Sample_Name == "Standard 95" ~ 28,
    Sample_Name == "Standard 96" ~ 32,
    Sample_Name == "Standard 97" ~ 40,
    TRUE ~ `Conc`),  # leave unchanged if no match
    Run_Date = format(lubridate::ymd_hms(Run_Time), "%m-%d-%Y"))

#ID x and y
x4 <- stds_SO4$Absorbance
y4 <- stds_SO4$Conc

stds_NO2 <- stds_NO2 %>%
  mutate(`Conc` = case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "1ppb NO2-N" ~ 0.89,
    Sample_Name == "2ppb NO2-N" ~ 1.78,
    Sample_Name == "5ppb NO2-N" ~ 4.89,
    Sample_Name == "10ppb NO2-N" ~ 10.22,
    Sample_Name == "25ppb NO2-N" ~ 24.89,
    Sample_Name == "50ppb NO2-N" ~ 50.22,
    Sample_Name == "75ppb NO2-N" ~ 75.11,
    Sample_Name == "150ppb NO2-N" ~ 150.22,
    Sample_Name == "200ppb NO2-N" ~ 200,
    TRUE ~ `Conc`),  # leave unchanged if no match
    Run_Date = format(lubridate::ymd_hms(Run_Time), "%m-%d-%Y"))

#ID x and y
x5 <- stds_NO2$Absorbance
y5 <- stds_NO2$Conc


#generating line of best fit aka standard curves
#NOx
lin_reg_NOx <- lm(y1 ~ x1)
NOx_coefs <- coef(lin_reg_NOx)                        # intercept and slope
NOx_r2 <- summary(lin_reg_NOx)$r.squared              # R squared
# Store in a clean dataframe
std_curve_NOx_1 <- data.frame(
  Test = "NOx",
  Intercept = NOx_coefs["(Intercept)"],
  Slope = NOx_coefs[2],
  R_squared = NOx_r2
)

#NH3
lin_reg_NH3 <- lm(y2 ~ x2)
NH3_coefs <- coef(lin_reg_NH3)                        # intercept and slope
NH3_r2 <- summary(lin_reg_NH3)$r.squared              # R squared
# Store in a clean dataframe
std_curve_NH3_1 <- data.frame(
  Test = "NH3",
  Intercept = NH3_coefs["(Intercept)"],
  Slope = NH3_coefs[2],
  R_squared = NH3_r2
)

#PO4
lin_reg_PO4 <- lm(y3 ~ x3)
PO4_coefs <- coef(lin_reg_PO4)                        # intercept and slope
PO4_r2 <- summary(lin_reg_PO4)$r.squared              # R squared
# Store in a clean dataframe
std_curve_PO4_1 <- data.frame(
  Test = "PO4",
  Intercept = PO4_coefs["(Intercept)"],
  Slope = PO4_coefs[2],
  R_squared = PO4_r2
) 

#SO4
# Fit the 3-polynomial model
triple_reg_SO4 <- lm(y1 ~ x1 + I(x1^2)+ I(x1^3))

SO4_coefs <- coef(triple_reg_SO4)                        # intercept and slope
SO4_r2 <- summary(triple_reg_SO4)$r.squared              # R squared
# Store in a clean dataframe
std_curve_SO4_1 <- data.frame(
  Test = "SO4",
  Intercept = SO4_coefs["(Intercept)"],
  Slope = SO4_coefs[2],
  R_squared = SO4_r2
) 

#NO2
lin_reg_NO2 <- lm(y5 ~ x5)
NO2_coefs <- coef(lin_reg_NO2)                        # intercept and slope
NO2_r2 <- summary(lin_reg_NO2)$r.squared              # R squared
# Store in a clean dataframe
std_curve_NO2_1 <- data.frame(
  Test = "NO2",
  Intercept = NO2_coefs["(Intercept)"],
  Slope = NO2_coefs[2],
  R_squared = NO2_r2
) 


#Combining standards & curve data frames
stds_combo <- bind_rows(stds_NOx, stds_NH3, stds_PO4, stds_SO4, stds_NO2)
std_curve_combo <- bind_rows(std_curve_NOx_1, std_curve_NH3_1, std_curve_PO4_1, std_curve_SO4_1, std_curve_NO2_1)


```

## Plot standard curves 
```{r Plot Standard Curves, echo=FALSE}

#Plot standard Curve or Curves 
  #v-Nox
std_curve_NOx <- ggplot(stds_NOx, aes(x = Conc, y = Absorbance)) +
  geom_point(size=3) +
  geom_smooth(method = "lm", se = FALSE, color = "coral") +
  theme_bw() + 
  facet_wrap(~Run_Date) +
  labs(x="Concentration (ppm)",
       y="Absorbance",
       title = "NOx Standard Curve")+ 
  annotate("text", x = 0.75, y = max(stds_NOx$Absorbance), 
           label = paste("r^2 =", round(NOx_r2, 4)),
           color = "black", size = 4)
print(std_curve_NOx)

  #NH3 
std_curve_NH3 <- ggplot(stds_NH3, aes(x = Conc, y = Absorbance)) +
  geom_point(size=3) +
  geom_smooth(method = "lm", se = FALSE, color = "darkgoldenrod") +
  theme_bw() + 
  facet_wrap(~Run_Date) +
  labs(x="Concentration (ppm)",
       y="Absorbance",
       title = "NH3 Standard Curve")+ 
  annotate("text", x = 1.7, y = max(stds_NH3$Absorbance), 
           label = paste("r^2 =", round(NH3_r2, 4)),
          color = "black", size = 4)
print(std_curve_NH3)

  #PO4 
std_curve_PO4 <- ggplot(stds_PO4, aes(x = Conc, y = Absorbance)) +
  geom_point(size=3) +
  geom_smooth(method = "lm", se = FALSE, color = "darkkhaki") +
  facet_wrap(~Run_Date) +
  theme_bw() + 
  labs(x="Concentration (ppm)",
       y="Absorbance",
       title = "PO4 Standard Curve")+ 
  annotate("text", x = 0.25, y = max(stds_PO4$Absorbance), label = paste("r^2 =", round(PO4_r2, 4)),
          color = "black", size = 4)
print(std_curve_PO4)

  #SO4 
std_curve_SO4 <- ggplot(stds_SO4, aes(x = Conc, y = Absorbance)) +
  geom_point(size=3) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3, raw = TRUE), se = FALSE, color = "darkgrey") +
  facet_wrap(~Run_Date) +
  theme_bw() + 
  labs(x="Concentration (ppm)",
       y="Absorbance",
       title = "SO4 Standard Curve")+ 
  annotate("text", x = 10, y = max(stds_SO4$Absorbance), label = paste("r^2 =", round(SO4_r2, 4)),
          color = "black", size = 4)
print(std_curve_SO4)

  #NO2 
std_curve_NO2 <- ggplot(stds_NO2, aes(x = Conc, y = Absorbance)) +
  geom_point(size=3) +
  geom_smooth(method = "lm", se = FALSE, color = "darkgreen") +
  facet_wrap(~Run_Date) +
  theme_bw() + 
  labs(x="Concentration (ppm)",
       y="Absorbance",
       title = "NO2 Standard Curve")+ 
  annotate("text", x = 10, y = max(stds_NO2$Absorbance), label = paste("r^2 =", round(NO2_r2, 4)),
          color = "black", size = 4)
print(std_curve_NO2)

############## Report on Cutoffs 

#Report out a flag if the run has an R2 lower than appropriate 
#Write out to the user whether or not the r2 is above the cutoff of 0.98
  ifelse(std_curve_NOx_1$R_squared <= r2_cutoff, 
         "NOx Curve r2 is below cutoff! - REASSESS", 
         "NOx Curve r2 GOOD - PROCEED")
  ifelse(std_curve_NH3_1$R_squared <= r2_cutoff, 
         "NH3 Curve r2 is below cutoff! - REASSESS", 
         "NH3 Curve r2 GOOD - PROCEED")
  ifelse(std_curve_PO4_1$R_squared <= r2_cutoff, 
         "PO4 Curve r2 is below cutoff! - REASSESS", 
         "PO4 Curve r2 GOOD - PROCEED")
  ifelse(std_curve_SO4_1$R_squared <= r2_cutoff, 
         "SO4 Curve r2 is below cutoff! - REASSESS", 
         "SO4 Curve r2 GOOD - PROCEED")
  ifelse(std_curve_NO2_1$R_squared <= r2_cutoff, 
         "NO2 Curve r2 is below cutoff! - REASSESS", 
         "NO2 Curve r2 GOOD - PROCEED")
  
  

#check for the a slope QAQC file, if there is not one, make one 
if (file.exists(file_path)) {
  # If it exists, read it back into R
  stds_log <- read.csv(file_path)
  print("QAQC log file exists and has been read into the code.")
  } else {
  # If it does not exist, create the CSV file
  stds_log <- as.data.frame(matrix(ncol = 6, nrow = 0))
  colnames(stds_log) <- c("Test", "Intercept", "Slope", "R_squared", "run_date", "user")
  
  #maybe add here the last slopes from the previous year
  
  # Write add_log to CSV
  write.csv(stds_log, file = file_path, row.names = FALSE)
  print("QAQC log file does not exist. It has been created.")
  }
  
  #getting rid of first "X" column
  #stds_log <- stds_log %>%
  #  select(-1)
  
#create a dataframe that can be appended to the QAQC log 
add_log <- as.data.frame(bind_rows(std_curve_NOx_1, std_curve_NH3_1, std_curve_PO4_1, std_curve_SO4_1, std_curve_NO2_1))
add_log$run_date <- run_date
add_log$user <- user

stds_log$run_date <- as.character(stds_log$run_date)
add_log$run_date  <- as.character(add_log$run_date)     #getting the run_date column in the same type to merge

#Filter to only rows in Slopes that are NOT already in log (by run_date + analyte) but need to add if none
if (nrow(stds_log) == 0) {
  # If log is empty, everything in add_log is new
  new_rows <- add_log
} else {
  # Otherwise, compare to existing log
  new_rows <- anti_join(add_log, stds_log, by = c("run_date", "Test"))
}


# Append the new, non-duplicate rows to log
log <- rbind(stds_log, new_rows)


#Plot the slopes through time 
Slopes_chk <- ggplot(log, aes(run_date, Slope, col=Test)) + 
  geom_point(size=4) + 
  geom_line(aes(group = Test)) + 
  theme_bw() + 
  labs(title="Slope Drift Assessment", x="Run Date", y="Slope") +
  scale_color_manual(values=c("coral", "darkgoldenrod", "darkkhaki", "darkgrey", "darkgreen"))
Slopes_chk

#averaging the slopes from the log and printing
avg_slopes <- log %>%
  group_by(Test) %>%
  summarise(avg_slope = mean(Slope, na.rm = TRUE))
knitr::kable(avg_slopes, caption = "Average Slope by Analyte", digits = 3)


#write out the log file with the added lines for this run date 
write.csv(log, file_path)

#write out a flag to the sample dataframe if the r2 is above the cutoff of 0.98
df_all <- df_all %>%
  mutate(
    NOx_flag = if (NOx_r2 <= r2_cutoff) {
      "NOx r2 low"
    } else {
      ""
    },
    NH3_flag = if (NH3_r2 <= r2_cutoff) {
      "NH3 r2 low"
    } else {
      ""
    },
    PO4_flag = if (PO4_r2 <= r2_cutoff) {
      "PO4 r2 low"
    } else {
      ""
    }, 
    SO4_flag = if (SO4_r2 <= r2_cutoff) {
      "SO4 r2 low"
    } else {
      ""
    }, 
    NO2_flag = if (NO2_r2 <= r2_cutoff) {
      "NO2 r2 low"
    } else {
      ""
    }
  )



```
 
 \newpage


##Dilution Corrections - ensure the latest dilution is kept
```{r Dilution Corrections, echo=FALSE}

####It looks like all duplicates were already removed?####

#Calculate the concentration when accounting for the dilution factor if applicable

#checking to see if duplicate samples (mislabeled/wrongly inputted/reran)
df_duplicates <- df_all %>%
  filter(!if_all(everything(), is.na)) %>%  # Remove rows that are entirely NA

  # Exclude common QC sample names and standard patterns
  filter(
    !(Sample_Name %in% c(
      "Duplicate", "CCV", "CCB", "1mg/L ammonia", "DI",
      "Blank", "Auto Spike", "0.15 mg/L"
    )) &
    !startsWith(Sample_Name, "Standard ") &
    !startsWith(Sample_Name, "Nitrite ") &
    !startsWith(Sample_Name, "Nitrate ") &
    !startsWith(Sample_Name, "Abs_Chk_") &
    !startsWith(Sample_Name, "peChk_") &
    !grepl("^\\d{1,2}/\\d{1,2}/\\d{4}\\s+\\d{1,2}:\\d{2}$", Sample_Name) &  # Remove timestamp-like names
    !grepl("^\\d+$", Sample_Name) &  # Remove sample names that are just numbers
    !is.na(Sample_Name) & Sample_Name != ""  # ðŸ” Remove blank or missing names
  ) %>%
  group_by(Test, Sample_Name) %>%  # Group by test and sample
  filter(n() > 1) %>%              # Keep only those that appear more than once
  ungroup()

#pulling them out into a string and printing them out
dil_dups_string <- df_duplicates %>%
  distinct(Sample_Name) %>%
  pull(Sample_Name) %>%
  paste(collapse = ", ")

# Make sure Dilution is numeric
df_duplicates$Dilution <- as.numeric(df_duplicates$Dilution)

# Report duplicated samples
if (dil_dups_string != "") {
  message("Duplicated samples: ", dil_dups_string)

  bad_dups <- df_duplicates %>%
    group_by(Test, Sample_Name) %>%
    summarise(
      n_dups = n(),
      any_dilution = any(Dilution > 1, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    filter(!any_dilution)

  if (nrow(bad_dups) > 0) {
    cat("\nWARNING: Duplicated samples with NO dilution (possible input error):\n")
    print(bad_dups)
  } else {
    cat("\n All duplicated samples have valid dilutions. No naming issues detected.\n")
  }

} else {
  message("No duplicated samples.")
}


df_all_cor <- df_all %>%
  arrange(row_number()) %>%
  mutate(
    Dilution = case_when(
      Dilution == "0" ~ 1.0,
      TRUE ~ as.numeric(Dilution)
    ),
    Pair_ID = ifelse(
      grepl("Duplicate", Sample_Name, ignore.case = TRUE),
      paste0(lag(Sample_Name), "_", lag(Run_Number)),  # pair only if same run
      paste0(Sample_Name, "_", Run_Number)
    ),
    Original_Name = Sample_Name
  ) %>%
  group_by(Pair_ID, Test, Run_Number) %>%  # group by Pair_ID, Test, and Run_Number
  filter(
    Dilution == ifelse(any(Dilution > 1, na.rm = TRUE), max(Dilution, na.rm = TRUE), 1)
  ) %>%
  ungroup()


```

 
## Check NOx Reduction Efficiency 
```{r Assess reduction Efficiency, echo=FALSE}

#check on the reduction efficiency for the NOx test: 
#Pull out test stds 
red_eff <- df_all %>%
  filter(Sample_Name %in% c("Nitrate Standard", "Nitrite Standard"))

red_eff$Eff_Percent <- (red_eff$Conc / 1)*100 ####Should we be dividing by 1 for this? These standards are 1 ppm NO3 and NO2. ####

red_eff$Eff_flag <-  ifelse(red_eff$Eff_Percent >= 90, 'YES', 'NO, rerun')

red_eff <- red_eff %>%
  mutate(row_num = row_number())

#Plot these efficencies 
red_effs_plot <-  ggplot(data = red_eff, aes(x = factor(row_num), y = Conc, fill=Eff_flag)) +
       geom_bar(stat = 'identity') + 
       facet_wrap(~Sample_Name) +
       scale_fill_manual(values = c("YES" = "midnightblue", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NOx (mg/L)", title=" ") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept= 1, linetype="dashed", 
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Reduction Efficiency >90%"))+
      scale_x_discrete(drop = TRUE)
print(red_effs_plot)

#create a flag 
NOx_mean_red_eff <- mean(red_eff$Eff_Percent)

#report out if flags indicate need for rerun and what the value is
ifelse(NOx_mean_red_eff >= 95,
       "Mean NOx Reduction Efficiency >95% - PROCEED",
       "Mean NOx Reduction Efficiency <95% - REASSESS")
print(NOx_mean_red_eff)

#write out a flag to the sample dataframe if red eff is low
if (NOx_mean_red_eff <= 95) {
    df_all$NOx_flag <- ifelse(
    df_all$NOx_flag != "",
    paste0(df_all$NOx_flag, "; NOx reduction efficiency low"),
    "NOx reduction efficiency low")}

```

\newpage

## Analyze the Check Standards
```{r Check Standards, echo=FALSE}

#Pull out check standards from raw file 
CCV_chks <- df_all %>%
  filter(str_detect(Sample_Name, "CCV"))

NOx_chks <- CCV_chks %>%
  filter(str_detect(Test, "Nitrate"))%>% 
  mutate(rep = row_number())

NH3_chks <- CCV_chks %>%
  filter(str_detect(Test, "Ammonia"))%>% 
  mutate(rep = row_number())

PO4_chks <- CCV_chks %>%
  filter(str_detect(Test, "Phosphate"))%>% 
  mutate(rep = row_number())

SO4_chks <- CCV_chks %>%
  filter(str_detect(Test, "Sulfate"))%>% 
  mutate(rep = row_number())

NO2_chks <- CCV_chks %>%
  filter(str_detect(Test, "Nitrite"))%>% 
  mutate(rep = row_number())

#RSV of standards 
chks_NOx_rsv <- ((sd(NOx_chks$Conc))/mean(NOx_chks$Conc))
chks_NH3_rsv <- ((sd(NH3_chks$Conc))/mean(NH3_chks$Conc))
chks_PO4_rsv <- ((sd(PO4_chks$Conc))/mean(PO4_chks$Conc))
chks_SO4_rsv <- ((sd(SO4_chks$Conc))/mean(SO4_chks$Conc))
chks_NO2_rsv <- ((sd(NO2_chks$Conc))/mean(NO2_chks$Conc))

#write out to user about whether or not to continue 
ifelse(chks_NOx_rsv >= chk_flag, 
       "NOx CHECK STANDARD RSD TOO HIGH - REASSESS",
       "NOx Check Standard RSD within Range - PROCEED")
ifelse(chks_NH3_rsv >= chk_flag, 
       "NH3 CHECK STANDARD RSD TOO HIGH - REASSESS",
       "NH3 Check Standard RSD within Range - PROCEED")
ifelse(chks_PO4_rsv >= chk_flag, 
       "PO4 CHECK STANDARD RSD TOO HIGH - REASSESS",
       "PO4 Check Standard RSD within Range - PROCEED")
ifelse(chks_SO4_rsv >= chk_flag, 
       "SO4 CHECK STANDARD RSD TOO HIGH - REASSESS",
       "SO4 Check Standard RSD within Range - PROCEED")
ifelse(chks_NO2_rsv >= chk_flag, 
       "NO2 CHECK STANDARD RSD TOO HIGH - REASSESS",
       "NO2 Check Standard RSD within Range - PROCEED")

#calculate percent difference between check standards & expected concentration 
#flag if the percent difference is over X% (defined in setup)
#calculate percent difference of check standards 
NOx_chks$NOx_diff <- (abs(NOx_chks$Conc - NOx_CCV)/((NOx_chks$Conc + NOx_CCV)/2)) * 100
NOx_chks$NOx_diff_flag <-  ifelse(NOx_chks$NOx_diff <= chk_conc_flag, 'YES', 'NO, rerun')

NH3_chks$NH3_diff <- (abs(NH3_chks$Conc - NH3_CCV)/((NH3_chks$Conc + NH3_CCV)/2)) * 100
NH3_chks$NH3_diff_flag <-  ifelse(NH3_chks$NH3_diff <= chk_conc_flag, 'YES', 'NO, rerun')

PO4_chks$PO4_diff <- (abs(PO4_chks$Conc - PO4_CCV)/((PO4_chks$Conc + PO4_CCV)/2)) * 100
PO4_chks$PO4_diff_flag <-  ifelse(PO4_chks$PO4_diff <= chk_conc_flag, 'YES', 'NO, rerun')

SO4_chks$SO4_diff <- (abs(SO4_chks$Conc - SO4_CCV)/((SO4_chks$Conc + SO4_CCV)/2)) * 100
SO4_chks$SO4_diff_flag <-  ifelse(SO4_chks$SO4_diff <= chk_conc_flag, 'YES', 'NO, rerun')

NO2_chks$NO2_diff <- (abs(NO2_chks$Conc - NO2_CCV)/((NO2_chks$Conc + NO2_CCV)/2)) * 100
NO2_chks$NO2_diff_flag <-  ifelse(NO2_chks$NO2_diff <= chk_conc_flag, 'YES', 'NO, rerun')


#plot the check standards
NOx_chks_plot <-  ggplot(data = NOx_chks, aes(x = rep, y = Conc, fill=NOx_diff_flag)) +
       geom_bar(stat = 'identity') + 
        scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
        labs(x= " ", y="NOx (mg/L)", title="Check Stds: NOx") + 
        theme(legend.position="bottom") +  
        geom_hline(yintercept=NOx_CCV,
              linetype="dashed", color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="% Difference <10%")) +
        theme_classic()

NH3_chks_plot <-  ggplot(data = NH3_chks, aes(x = rep, y = Conc, fill=NH3_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
       theme_classic() + 
       labs(x= " ", y="NH3  (mg/L)", title="Check Stds: NH3") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept=NH3_CCV,
              linetype="dashed",  color = "black", linewidth=1) + 
       guides(fill=guide_legend(title="% Difference <10%"))

PO4_chks_plot <-  ggplot(data = PO4_chks, aes(x = rep, y = Conc, fill=PO4_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
        theme_classic() + labs(x= " ", y="PO4  (mg/L)", title="Check Stds: PO4") + 
        theme(legend.position="bottom") +  
        geom_hline(yintercept=PO4_CCV,
              linetype="dashed",  color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="% Difference <10%"))

SO4_chks_plot <-  ggplot(data = SO4_chks, aes(x = rep, y = Conc, fill=SO4_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
        theme_classic() + labs(x= " ", y="SO4  (mg/L)", title="Check Stds: SO4") + 
        theme(legend.position="bottom") +  
        geom_hline(yintercept=SO4_CCV,
              linetype="dashed",  color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="% Difference <10%"))

NO2_chks_plot <-  ggplot(data = NO2_chks, aes(x = rep, y = Conc, fill=NO2_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
        theme_classic() + labs(x= " ", y="NO2  (mg/L)", title="Check Stds: NO2") + 
        theme(legend.position="bottom") +  
        geom_hline(yintercept=NO2_CCV,
              linetype="dashed",  color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="% Difference <10%"))

#Combined Plot
Chk_Std_Plots <- ggarrange(
  NOx_chks_plot, 
  NH3_chks_plot, 
  PO4_chks_plot, 
  SO4_chks_plot,   
  NO2_chks_plot,   
  nrow = 2, 
  ncol = 3,
  common.legend = TRUE,
  legend = "bottom")
print(Chk_Std_Plots)

#calculate the percent of check standards that are within the range based on the flag 
NOx_chks_percent <- (sum(NOx_chks$NOx_diff_flag == "YES")/nrow(NOx_chks))*100
NH3_chks_percent <- (sum(NH3_chks$NH3_diff_flag == "YES")/nrow(NH3_chks))*100
PO4_chks_percent <- (sum(PO4_chks$PO4_diff_flag == "YES")/nrow(PO4_chks))*100
SO4_chks_percent <- (sum(SO4_chks$PO4_diff_flag == "YES")/nrow(SO4_chks))*100
NO2_chks_percent <- (sum(NO2_chks$PO4_diff_flag == "YES")/nrow(NO2_chks))*100


#report out if flags indicate need for rerun
ifelse(NOx_chks_percent >= chks_flag, 
       ">60% of NOx Check Standards are within range of expected concentration - PROCEED",
       "<60% of NOx Check Standards are within range of expected concentration - REASSESS")
ifelse(NH3_chks_percent >= chks_flag,
       ">60% of NH3 Check Standards are within range of expected concentration - PROCEED",
       "<60% of NH3 Check Standards are within range of expected concentration - REASSESS")
ifelse(PO4_chks_percent >= chks_flag,
       ">60% of PO4 Check Standards are within range of expected concentration - PROCEED",
       "<60% of PO4 Check Standards are within range of expected concentration - REASSESS")
ifelse(SO4_chks_percent >= chks_flag,
       ">60% of SO4 Check Standards are within range of expected concentration - PROCEED",
       "<60% of SO4 Check Standards are within range of expected concentration - REASSESS")
ifelse(NO2_chks_percent >= chks_flag,
       ">60% of NO2 Check Standards are within range of expected concentration - PROCEED",
       "<60% of NO2 Check Standards are within range of expected concentration - REASSESS")


#write out a flag to the sample dataframe if less than 60% of the checks are within the expected CV
if (NOx_chks_percent <= chks_flag) {
    df_all_cor$NOx_flag <- ifelse(
    df_all_cor$NOx_flag != "",
    paste0(df_all_cor$NOx_flag, "; NOx checks out of range"),
    "NOx checks out of range"
  )
}

if (NH3_chks_percent <= chks_flag) {  
    df_all_cor$NH3_flag <- ifelse(
    df_all_cor$NH3_flag != "",
    paste0(df_all_cor$NH3_flag, "; NH3 checks out of range"),
    "NH3 checks out of range"
  )
}

if (PO4_chks_percent <= chks_flag) {  
    df_all_cor$PO4_flag <- ifelse(
    df_all_cor$PO4_flag != "",
    paste0(df_all_cor$PO4_flag, "; PO4 checks out of range"),
    "PO4 checks out of range"
  )
}

if (SO4_chks_percent <= chks_flag) {  
    df_all_cor$SO4_flag <- ifelse(
    df_all_cor$SO4_flag != "",
    paste0(df_all_cor$SO4_flag, "; SO4 checks out of range"),
    "SO4 checks out of range"
  )
}

if (NO2_chks_percent <= chks_flag) {  
    df_all_cor$NO2_flag <- ifelse(
    df_all_cor$NO2_flag != "",
    paste0(df_all_cor$NO2_flag, "; NO2 checks out of range"),
    "NO2 checks out of range"
  )
}


```

\newpage

## Analyze Blanks 
```{r Analyze Blanks, echo=FALSE}

## CCB was used as the blank for all analytes

#Pull out check standards from raw file 
NOx_blks <- df_all %>%
  filter(Test == "Nitrate (NOx)")%>% 
  filter(str_detect(Sample_Name, "CCB"))%>% 
  mutate(rep = row_number())

NH3_blks <- df_all %>%
  filter(Test == "Ammonia (PPS)")%>% 
  filter(str_detect(Sample_Name, "CCB"))%>% 
  mutate(rep = row_number())

PO4_blks <- df_all %>%
  filter(Test == "Phosphate ")%>% 
  filter(str_detect(Sample_Name, "CCB"))%>% 
  mutate(rep = row_number())

SO4_blks <- df_all %>%
  filter(Test == "Sulfate")%>% 
  filter(str_detect(Sample_Name, "CCB"))%>% 
  mutate(rep = row_number())

NO2_blks <- df_all %>%
  filter(Test == "Nitrite")%>% 
  filter(str_detect(Sample_Name, "CCB"))%>% 
  mutate(rep = row_number())

#Pull out samples from df_all to calc quantile
samples <- df_all %>%  
  filter(str_detect(Sample_Name, c("cm"))) ####Not sure if this is the best way to identify samples####

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations
NOx_only <- samples %>%
  filter(Test == "Nitrate (NOx)")
blk_flag_NOx <- quantile(NOx_only$Conc, prob=c(.25))   #this gives you the lower 25% quartile of the data 
blk_flag_NOx <- ifelse(blk_flag_NOx < NOx_dl, NOx_dl/2, blk_flag_NOx) #this will determine if the Q1 of samples is below the detection limit, if it is then it will replace the flag with half of the detection limit 
NOx_blks$NOx_diff_flag <-  ifelse(NOx_blks$Conc <= blk_flag_NOx, 'YES', 'NO, rerun')

NH3_only <- samples %>%
  filter(Test == "Ammonia (PPS)")
blk_flag_NH3 <- quantile(NH3_only$Conc, prob=c(.25))   #this gives you the lower 25% quartile of the data 
blk_flag_NH3 <- ifelse(blk_flag_NH3 < NH3_dl, NH3_dl/2, blk_flag_NH3) #this will determine if the Q1 of samples is below the detection limit, if it is then it will replace the flag with half of the detection limit 
NH3_blks$NH3_diff_flag <-  ifelse(NH3_blks$Conc <= blk_flag_NH3, 'YES', 'NO, rerun')

PO4_only <- samples %>%
  filter(Test == "Phosphate ")
blk_flag_PO4 <- quantile(PO4_only$Conc, prob=c(.25))   #this gives you the lower 25% quartile of the data 
blk_flag_PO4 <- ifelse(blk_flag_PO4 < PO4_dl, PO4_dl/2, blk_flag_PO4) #this will determine if the Q1 of samples is below the detection limit, if it is then it will replace the flag with half of the detection limit 
PO4_blks$PO4_diff_flag <-  ifelse(PO4_blks$Conc <= blk_flag_PO4, 'YES', 'NO, rerun')

SO4_only <- samples %>%
  filter(Test == "Sulfate")
blk_flag_SO4 <- quantile(SO4_only$Conc, prob=c(.25))   #this gives you the lower 25% quartile of the data 
blk_flag_SO4 <- ifelse(blk_flag_SO4 < SO4_dl, SO4_dl/2, blk_flag_SO4) #this will determine if the Q1 of samples is below the detection limit, if it is then it will replace the flag with half of the detection limit 
SO4_blks$SO4_diff_flag <-  ifelse(SO4_blks$Conc <= blk_flag_SO4, 'YES', 'NO, rerun')

NO2_only <- samples %>%
  filter(Test == "Nitrite")
blk_flag_NO2 <- quantile(NO2_only$Conc, prob=c(.25))   #this gives you the lower 25% quartile of the data 
blk_flag_NO2 <- ifelse(blk_flag_NO2 < NO2_dl, NO2_dl/2, blk_flag_NO2) #this will determine if the Q1 of samples is below the detection limit, if it is then it will replace the flag with half of the detection limit 
NO2_blks$NO2_diff_flag <-  ifelse(NO2_blks$Conc <= blk_flag_NO2, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag 
NOx_blks_percent <- (sum(NOx_blks$NOx_diff_flag == "YES")/nrow(NOx_blks))*100
NH3_blks_percent <- (sum(NH3_blks$NH3_diff_flag == "YES")/nrow(NH3_blks))*100
PO4_blks_percent <- (sum(PO4_blks$PO4_diff_flag == "YES")/nrow(PO4_blks))*100
SO4_blks_percent <- (sum(SO4_blks$SO4_diff_flag == "YES")/nrow(SO4_blks))*100
NO2_blks_percent <- (sum(NO2_blks$NO2_diff_flag == "YES")/nrow(NO2_blks))*100

#report out if flags indicate need for rerun
ifelse(NOx_blks_percent >= 60,
       ">60% of NOx Blanks are below the lower 25% quartile of samples or 1/2 detection limit - PROCEED",
       "<60% of NOx Blanks are lower 25% quartile of samples or 1/2 the detection limit - REASSESS")
ifelse(NH3_blks_percent >= 60,
       ">60% of NH3 Blanks are below the lower 25% quartile of samples - PROCEED",
       "<60% of NH3 Blanks are lower 25% quartile of samples - REASSESS")
ifelse(PO4_blks_percent >= 60,
       ">60% of PO4 Blanks  are below the lower 25% quartile of samples- PROCEED",
       "<60% of PO4 Blanks are lower 25% quartile of samples - REASSESS")
ifelse(SO4_blks_percent >= 60,
       ">60% of SO4 Blanks  are below the lower 25% quartile of samples- PROCEED",
       "<60% of SO4 Blanks are lower 25% quartile of samples - REASSESS")
ifelse(NO2_blks_percent >= 60,
       ">60% of NO2 Blanks  are below the lower 25% quartile of samples- PROCEED",
       "<60% of NO2 Blanks are lower 25% quartile of samples - REASSESS")

#plotting the blanks compared to the lower 25% of conc (what the flag is)
NOx_blks_plot <-  ggplot(data = NOx_blks, aes(x = rep, y = Conc, fill=NOx_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NOx (mg/L)", title="NOx Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept= as.numeric(blk_flag_NOx), linetype="dashed", 
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))

NH3_blks_plot <-  ggplot(data = NH3_blks, aes(x = rep, y = Conc, fill=NH3_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NH3  (mg/L)", title="NH3 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept=as.numeric(blk_flag_NH3), linetype="dashed",
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))

PO4_blks_plot <-  ggplot(data = PO4_blks, aes(x = rep, y = Conc, fill=PO4_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="PO4 (mg/L)", title="PO4 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept=as.numeric(blk_flag_PO4), linetype="dashed", 
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))

SO4_blks_plot <-  ggplot(data = SO4_blks, aes(x = rep, y = Conc, fill=SO4_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="SO4 (mg/L)", title="SO4 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept=as.numeric(blk_flag_SO4), linetype="dashed", 
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))

NO2_blks_plot <-  ggplot(data = NO2_blks, aes(x = rep, y = Conc, fill=NO2_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NO2 (mg/L)", title="NO2 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept=as.numeric(blk_flag_NO2), linetype="dashed", 
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))


#Combined Plot
Blks_Plots <- ggarrange(
  NOx_blks_plot, 
  NH3_blks_plot, 
  PO4_blks_plot,
  SO4_blks_plot,
  NO2_blks_plot,
  nrow = 2, 
  ncol = 3, 
  common.legend = TRUE,
  legend = "bottom")
print(Blks_Plots)

#find average of run blanks for flagging samples later 
# Compute averages
blk_avg_NOx <- mean(NOx_blks$Conc, na.rm = TRUE)
blk_avg_NH3 <- mean(NH3_blks$Conc, na.rm = TRUE)
blk_avg_PO4 <- mean(PO4_blks$Conc, na.rm = TRUE)
blk_avg_SO4 <- mean(SO4_blks$Conc, na.rm = TRUE)
blk_avg_NO2 <- mean(NO2_blks$Conc, na.rm = TRUE)

# Create a data frame
blank_avgs <- data.frame(
  Test = c("NOx", "NH3", "PO4", "SO4", "NO2"),
  Blank_Mean_Conc = c(blk_avg_NOx, blk_avg_NH3, blk_avg_PO4, blk_avg_SO4, blk_avg_NO2)
)

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Concentration of Blanks", digits = 4)


#write out a flag to the sample dataframe if more than 60% of the blanks are above the lower 25% quantile of samples
if (NOx_blks_percent <= chks_flag) {
    df_all_cor$NOx_flag <- ifelse(
    df_all_cor$NOx_flag != "",
    paste0(df_all_cor$NOx_flag, "; NOx blanks out of range"),
    "NOx blanks out of range"
  )
}

if (NH3_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
    df_all_cor$NH3_flag <- ifelse(
    df_all_cor$NH3_flag != "",
    paste0(df_all_cor$NH3_flag, "; NH3 blanks out of range"),
    "NH3 blanks out of range"
  )
}

if (PO4_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
    df_all_cor$PO4_flag <- ifelse(
    df_all_cor$PO4_flag != "",
    paste0(df_all_cor$PO4_flag, "; PO4 blanks out of range"),
    "PO4 blanks out of range"
  )
}

if (SO4_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
    df_all_cor$SO4_flag <- ifelse(
    df_all_cor$SO4_flag != "",
    paste0(df_all_cor$SO4_flag, "; PO4 blanks out of range"),
    "PO4 blanks out of range"
  )
}

if (NO2_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
    df_all_cor$NO2_flag <- ifelse(
    df_all_cor$NO2_flag != "",
    paste0(df_all_cor$NO2_flag, "; PO4 blanks out of range"),
    "PO4 blanks out of range"
  )
}
```

\newpage



```{r Unit conversion to uM, include=FALSE}

#Convert values based on the Test ID
####Need to check this####
##Everything is mg/L except Nitrite which is ug/L
# ((mg/L) /1000) = g/L
# (g/L) / (g/mol)) = mol/L
# (mol/L) * 1,000,000 = umol/L

##Nitrite
# ((ug/L) /100,000) = g/L
# (g/L) / (g/mol)) = mol/L
# (mol/L) * 100,000 = umol/L

samples <- samples %>%
  mutate(
    Conc_uM = case_when(
      Test == "Nitrate (NOx)" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "Ammonia (PPS)" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "Phosphate " ~ (((as.numeric(samples$Conc))/Con1)/P_mw)*Con2,
      Test == "Sulfate" ~ (((as.numeric(samples$Conc))/Con1)/S_mw)*Con2,
      Test == "Nitrite" ~ (((as.numeric(samples$Conc))/Con2)/N_mw)*Con2,

      TRUE ~ as.numeric(NA)  # fallback in case of unexpected value
    ),
    # Replace negatives with 0 in 'value' and 'value_converted'
    Conc = pmax(Conc, 0),
    Conc_uM = pmax(Conc_uM, 0)
  )

head(samples)
```


##Sample Flagging - Within range of standard curve
```{r Sample Flagging, echo=FALSE, include = FALSE}

####Need to know detection limit for all analytes####

cat("Sample Flagging")

#Flagging data if the concentration is outside the standards range 
samples_flagged <- samples %>%
  mutate(
    Conc_flag = case_when(
      Test == "Nitrate (NOx)" & Conc < NOx_dl ~ "bdl",
      Test == "Nitrate (NOx)" & Conc > NOx_top    ~ "adl",
      Test == "Nitrate (NOx)"               ~ "Within_Range",
      
      Test == "Ammonia (PPS)" & Conc < NH3_dl ~ "bdl",
      Test == "Ammonia (PPS)" & Conc > NH3_top    ~ "adl",
      Test == "Ammonia (PPS)"                ~ "Within_Range",
      
      Test == "Phosphate " & Conc < PO4_dl  ~ "bdl",
      Test == "Phosphate " & Conc > PO4_top   ~ "adl",
      Test == "Phosphate "                ~ "Within_Range",
      
      Test == "Sulfate" & Conc < SO4_dl  ~ "bdl",
      Test == "Sulfate" & Conc > SO4_top   ~ "adl",
      Test == "Sulfate"                ~ "Within_Range",
      
      Test == "Nitrite" & Conc < NO2_dl  ~ "bdl",
      Test == "Nitrite" & Conc > NO2_top   ~ "adl",
      Test == "Nitrite"                ~ "Within_Range",
      
      TRUE ~ NA_character_  # fallback for unexpected values
    )
  )

```

##Pull out sample id information 
```{r Sample Identification, echo=FALSE, include = FALSE}

cat("Sample Processing")

# 1_43_R76_5292024_ch6_trE_cm
head(nutr_metadata)
head(samples_flagged)

samples_flagged <- samples_flagged %>%
  separate(
    col = Sample_Name,
    sep = "_",
    into = c("Box ", "Vial_number", "Vial_ID", "Date", "Chamber_ID", "Treatment", "Depth_Units"),
    remove = FALSE)

samples_flagged$Date <- paste0("0", samples_flagged$Date)
samples_flagged$Date_formatted <- as.Date(samples_flagged$Date, format = "%m%d%Y")
samples_flagged$Date_formatted <- format(samples_flagged$Date_formatted, format = "%Y%m%d")

samples_flagged_selected <- samples_flagged %>%
  rename(
    SEAL_Sample_Name = Sample_Name,
    Old_Date = Date, 
    Date = Date_formatted)  %>%
  select(SEAL_Sample_Name, 
         Vial_ID, 
         Run_Number, 
         Conc, 
         Absorbance, 
         Dilution, 
         Unit, 
         Test, 
         Run_Time, 
         NOx_flag, 
         NH3_flag, 
         PO4_flag, 
         SO4_flag, 
         NO2_flag, 
         Conc_uM, 
         Conc_flag)

```

## Check to see if samples run match metadata & merge info
```{r check sample ids with metadata, echo=FALSE}

cat("Check Sample IDs with Metadata")

#check to see if all samples are present in the metadata 
all_present <- all(samples_flagged_selected$Vial_ID %in% nutr_metadata$Vial_ID)

if (all_present) {
  message("All sample IDs are present in metadata.")
} else {
  message("Some sample IDs are missing from metadata.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(samples_flagged_selected$Sample_Name, nutr_metadata$NUTR_ID)
  print(missing_ids)
}

#merge metadata with sample run data 
merged_data <- samples_flagged_selected %>%
  left_join(nutr_metadata, by = "Vial_ID")


```

\newpage

## Visualize data
```{r Visualize data, echo=FALSE, fig.height=10, fig.width=15}
cat("Visualize Data")

### Nitrite + Nitrate 

NOx_forplot <- merged_data %>%
  filter(Test == "Nitrate (NOx)")

#group the data for plotting
NOx_forplot <- NOx_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_NOx_plot <-  ggplot(data = NOx_forplot, aes(x = Vial_ID, y = Conc, fill=Treatment)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
                 color="black") +
       facet_grid(~ Treatment, scales="free") +
        theme_classic() + 
        labs(x= " ", y="NOx (mg/L)", title="Porewater NOx") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())

viz_NOx_plot


### Ammonia 
NH3_forplot <- merged_data %>%
  filter(Test == "Ammonia (PPS)")

#group the data for plotting
NH3_forplot <- NH3_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_NH3_plot <-  ggplot(data = NH3_forplot, aes(x = Vial_ID, y = Conc, fill=Treatment)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
                 color="black") +
       facet_grid(~ Treatment, scales="free") +
        theme_classic() + 
        labs(x= " ", y="NH3 (mg/L)", title="Porewater NH3") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())

viz_NH3_plot


### Phosphate 
PO4_forplot <- merged_data %>%
  filter(Test == "Phosphate ")

#group the data for plotting
PO4_forplot <- PO4_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_PO4_plot <-  ggplot(data = PO4_forplot, aes(x = Vial_ID, y = Conc, fill=Treatment)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
                 color="black") +
       facet_grid(~ Treatment, scales="free") +
        theme_classic() + 
        labs(x= " ", y="PO4 (mg/L)", title="Porewater PO4") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())

viz_PO4_plot

### Sulfate 
SO4_forplot <- merged_data %>%
  filter(Test == "Sulfate")

#group the data for plotting
SO4_forplot <- SO4_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_SO4_plot <-  ggplot(data = SO4_forplot, aes(x = Vial_ID, y = Conc, fill=Treatment)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
                 color="black") +
       facet_grid(~ Treatment, scales="free") +
        theme_classic() + 
        labs(x= " ", y="SO4 (mg/L)", title="Porewater SO4") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())

viz_SO4_plot

### Nitrite 
NO2_forplot <- merged_data %>%
  filter(Test == "Nitrite")

#group the data for plotting
NO2_forplot <- NO2_forplot %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_NO2_plot <-  ggplot(data = NO2_forplot, aes(x = Vial_ID, y = Conc, fill=Treatment)) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
                 color="black") +
       facet_grid(~ Treatment, scales="free") +
        theme_classic() + 
        labs(x= " ", y="NO2 (mg/L)", title="Porewater NO2") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5)) + 
        theme_classic(base_size = 30) +
        theme(axis.text.x = element_blank())

viz_NO2_plot



```


##Export Processed Data
```{r Export Processed Data, include=FALSE}

cat("Export Processed Data")

#pivot the data set wider to make it wide format 
final_data1 <- merged_data %>%
  pivot_wider(
    id_cols = Vial_ID,
    names_from = Test,
    values_from = Conc,
    names_glue = "{Test}_Conc"
  )

final_data2 <- merged_data %>%
  pivot_wider(
    id_cols = Vial_ID,
    names_from = Test,
    values_from = Conc_uM,
    names_glue = "{Test}_Conc_uM"
  )

final_data3 <- merged_data %>%
  pivot_wider(
    id_cols = Vial_ID,
    names_from = Test,
    values_from = Conc_flag,
    names_glue = "{Test}_Conc_flag"
  )

head(merged_data)
#take out only the columns we want to merge 
df_all_clean_cols <- merged_data %>%
  select(Vial_ID, SEAL_Sample_Name, Box, Vial_number, Chamber_ID , Treatment , Depth_cm ,
         Year, Month, Day, Analysis.day..FAUSTO...JADE.LEC., 
         NOx_flag, NH3_flag, PO4_flag, SO4_flag, NO2_flag, 
         Notes)

df_all_clean_cols_one_row <- df_all_clean_cols %>%
  group_by(Vial_ID) %>%
  slice(1) %>%  # or use summarise() if you want to aggregate
  ungroup()

#merge these together
data_list <- list(df_all_clean_cols_one_row, final_data1, final_data2, final_data3)

final_data4 <- reduce(data_list, full_join, by = c("Vial_ID"))


#Add project information 
final_data_labeled <- final_data4 %>% 
  mutate(
    Project = "COMPASS: OWC Root Exclusion",   # new column with same value on every row
    Region = "LE",
    Run_notes = run_notes, 
    Analysis_rundate = print(run_date)# new column with notes about the run
  ) 

#Prepare data to be exported 
final_data <- final_data_labeled %>%
    rename(
    Analysis_day_FAUSTO_JADE_LEC = Analysis.day..FAUSTO...JADE.LEC., 
    
    NOx_Conc_mgL = `Nitrate (NOx)_Conc`,
    NOx_Conc_uM = `Nitrate (NOx)_Conc_uM`,
    NOx_Conc_Flag = `Nitrate (NOx)_Conc_flag`,
    NOx_QAQC_Flag = NOx_flag,
    
    NH3_Conc_mgL = `Ammonia (PPS)_Conc`,
    NH3_Conc_uM = `Ammonia (PPS)_Conc_uM`,
    NH3_Conc_Flag =`Ammonia (PPS)_Conc_flag`,
    NH3_QAQC_Flag = NH3_flag,
    
    PO4_Conc_mgL = `Phosphate _Conc`,
    PO4_Conc_uM = `Phosphate _Conc_uM`,
    PO4_Conc_Flag =`Phosphate _Conc_flag`,
    PO4_QAQC_Flag = PO4_flag,
    
    SO4_Conc_mgL = `Sulfate_Conc`,
    SO4_Conc_uM = `Sulfate_Conc_uM`,
    SO4_Conc_Flag =`Sulfate_Conc_flag`,
    SO4_QAQC_Flag = SO4_flag,
    
    NO2_Conc_mgL = `Nitrite_Conc`,
    NO2_Conc_uM = `Nitrite_Conc_uM`,
    NO2_Conc_Flag =`Nitrite_Conc_flag`,
    NO2_QAQC_Flag = NO2_flag,
    # add more rename pairs as needed
  ) %>%
  select(Project, Region, Vial_ID, SEAL_Sample_Name, Box, Vial_number, Chamber_ID , Treatment , Depth_cm ,
         Year, Month, Day, Analysis_day_FAUSTO_JADE_LEC,
         NOx_Conc_mgL, NOx_Conc_uM, NOx_Conc_Flag, NOx_QAQC_Flag,
         NH3_Conc_mgL, NH3_Conc_uM, NH3_Conc_Flag, NH3_QAQC_Flag,
         PO4_Conc_mgL, PO4_Conc_uM, PO4_Conc_Flag, PO4_QAQC_Flag,
         SO4_Conc_mgL, SO4_Conc_uM, SO4_Conc_Flag, SO4_QAQC_Flag,
         NO2_Conc_mgL, NO2_Conc_uM, NO2_Conc_Flag, NO2_QAQC_Flag,
         Analysis_rundate,  Run_notes, Notes)


#Write out data frame 
  write.csv(final_data, final_path)
  

```


#end



